Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-08 csf-2017/paper.tex
fix so document compiles
text about exec/enter cond

\section{A capability machine with local capabilities}
\begin{figure}
  \begin{align*}
    \addr   &\in& \Addrs &\isdef \nats\\
    w &\in&\Words &\isdef \ints + \Caps \\
    \perm   &\in& \Perms &::= \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid \exec\mid \entry\mid \rwx\mid \rwlx\\
    \gl&\in&\Globals & ::= \glob \mid \local \\
     && \Caps  &::= \left\{
                       \begin{multlined}
                         ((\perm,\gl),\start,\addrend,\addr)\mid\\
                         \start\in\Addrs,\addrend \in
                         \Addrs\cup \{\infty\}
                       \end{multlined} \right\}\\
    r       &\in& \RegName&\bnfdef \pcreg\mid r_0\mid r_1\mid\ldots\\
    \reg &\in& \Regs  &\isdef \RegName \rightarrow \Words\\
    m&\in& \Heaps &\isdef \Addrs \rightarrow \Words \\
    \Phi    &\in& \ExecConfs  &\isdef \Regs \times \Heaps \\
    &&\Confs &::= \ExecConfs + \{\failed \} + \{\halted\} \times \Heaps \\
    \ms     &\in& \MemSegments &::= \Addrs \parfun \Words
  \end{align*}

  \begin{array}{rcl}
    \lv    &::=& \refreg{r} \\
    \hv    &::=& \refheap{r}\\
    \rv    &::=& n \mid \lv \\
    i      &::=& 
                 \jmp{\lv} \mid 
                 \jnz{\lv}{\rv} \mid
                 \move{\lv}{\rv} \mid 
                 \load{\lv}{\hv} \mid 
                 \store{\hv}{\rv} \mid  \\
           &   & \plus{\lv}{\rv}{\rv} \mid 
                 \lea{\lv}{\rv} \mid 
                 \restricttwo{\lv}{\rv} \mid 
                 \subseg{\lv}{\rv}{\rv} \mid  \\
           &   & \isptr{\lv}{\rv} \mid 
                 \getp{\lv}{\lv} \mid 
                 \getl{\lv}{\lv} \mid 
                 \getb{\lv}{\lv} \mid
                 \gete{\lv}{\lv} \mid\\
           &   & \geta{\lv}{\lv} \mid 
                 \fail \mid
                 \halt 
  \end{array}
  \caption{The syntax of our capability machine assembly language.}
  \label{fig:syntax}
\end{figure}

@@

For capabilities with a write permission, $\writeCond{}$ must be
satisfied for the capabilities range of authority. Generally speaking,
write capabilities may be used to write any word, so the region that
governs the part of memory the write capability has authority over
must allow any capability safe word.
This is of course not entirely true because of the distinction between
write and permit-write-local capabilities. For write capabilities, the
region must require the contents of memory to be non-local and
capability safe words. For permit-write-local capabilities, the
non-local requirement is omitted.

The write condition uses $\var{localityReg}$ to pick an appropriate
region in the same way the read condition does.

The condition for the region to be address-stratified is there for
technical reasons.\lau{Dominique, you probably disagree with this, so
  please feel free to add something.}

>>>>

\section{A capability machine with local capabilities}
\begin{figure}
  \begin{align*}
    \addr   &\in& \Addrs &\isdef \nats\\
    w &\in&\Words &\isdef \ints + \Caps \\
    \perm   &\in& \Perms &::= \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid \exec\mid \entry\mid \rwx\mid \rwlx\\
    \gl&\in&\Globals & ::= \glob \mid \local \\
     && \Caps  &::= \left\{
                       \begin{multlined}
                         ((\perm,\gl),\start,\addrend,\addr)\mid\\
                         \start\in\Addrs,\addrend \in
                         \Addrs\cup \{\infty\}
                       \end{multlined} \right\}\\
    r       &\in& \RegName&\bnfdef \pcreg\mid r_0\mid r_1\mid\ldots\\
    \reg &\in& \Regs  &\isdef \RegName \rightarrow \Words\\
    m&\in& \Heaps &\isdef \Addrs \rightarrow \Words \\
    \Phi    &\in& \ExecConfs  &\isdef \Regs \times \Heaps \\
    &&\Confs &::= \ExecConfs + \{\failed \} + \{\halted\} \times \Heaps \\
    \ms     &\in& \MemSegments &::= \Addrs \parfun \Words
  \end{align*}
\[
  \begin{array}{rcl}
    \lv    &::=& \refreg{r} \\
    \hv    &::=& \refheap{r}\\
    \rv    &::=& n \mid \lv \\
    i      &::=& 
                 \jmp{\lv} \mid 
                 \jnz{\lv}{\rv} \mid
                 \move{\lv}{\rv} \mid 
                 \load{\lv}{\hv} \mid 
                 \store{\hv}{\rv} \mid  \\
           &   & \plus{\lv}{\rv}{\rv} \mid 
                 \lea{\lv}{\rv} \mid 
                 \restricttwo{\lv}{\rv} \mid 
                 \subseg{\lv}{\rv}{\rv} \mid  \\
           &   & \isptr{\lv}{\rv} \mid 
                 \getp{\lv}{\lv} \mid 
                 \getl{\lv}{\lv} \mid 
                 \getb{\lv}{\lv} \mid
                 \gete{\lv}{\lv} \mid\\
           &   & \geta{\lv}{\lv} \mid 
                 \fail \mid
                 \halt 
  \end{array}
\]
  \caption{The syntax of our capability machine assembly language.}
  \label{fig:syntax}
\end{figure}

@@

For capabilities with a write permission, $\writeCond{}$ must be
satisfied for the capabilities range of authority. Generally speaking,
write capabilities may be used to write any word, so the region that
governs the part of memory the write capability has authority over
must allow any capability safe word.
This is of course not entirely true because of the distinction between
write and permit-write-local capabilities. For write capabilities, the
region must require the contents of memory to be non-local and
capability safe words. For permit-write-local capabilities, the
non-local requirement is omitted.
The write condition uses $\var{localityReg}$ to pick an appropriate
region in the same way the read condition does.
The condition for the region to be address-stratified is there for
technical reasons.\lau{Dominique, you probably disagree with this, so
  please feel free to add something.}

The conditions $\entryCond{}$ and $\execCond{}$ are very similar and
differs only slightly. Executable capabilities can be updated to point
anywhere in memory, so they must be capability safe as a program
counter no matter where they point in memory. Contrarily, the enter
capabilities are opaque and must be used from the address they point
to. They change permission when jumped to, so we require them to be
capability safe as a program counter when the permission is changed to
read-execute.
In both conditions, the capability must be capability safe as a
program counter certain future worlds. If the capability is global,
then it should be capability safe as a program counter in any private
future world. This means that it can only depend on permanent regions
because temporary regions may be revoked in a private future
world. On the other hand, local capabilities must only be capability
safe in public future worlds because they may depend on temporary
regions, so we cannot expect them to function correctly as program
counters if temporary regions are revoked.
